\section{Acknowledgements}

As I reach this significant milestone in my academic and personal
journey, I feel compelled to pause and honor those who have served as
irreplaceable pillars of support and inspiration.

Firstly, my deepest gratitude goes to my family. To my son, António,
your arrival in the middle of my Master\textquotesingle s program has
been the most beautiful and motivating challenge of my life. You are a
constant source of inspiration, a daily reminder of why I sought this
new path. To Ricardo, your love has been an unwavering presence,
fundamentally shaping how we\textquotesingle ve faced the challenges and
triumphs in our lives. To my parents, your support and love helped me
refine my commitment. Thank you for challenging me and for supporting my
choices. To my siblings, who have been invaluable in providing support
and in creating a sense of home, no matter the distance that separates
us.

My friends, who have been unwavering champions of my aspirations,
continually push me to break boundaries. Thank you for all the joy and
fulfillment you bring into my life.

I reserve a profound sense of gratitude for my academic advisors,
Professor Tiago Taveira-Gomes and Professor Ricardo Cruz-Correia. It was
under your mentorship that I discovered a new career path where my
clinical experience can merge with technological innovation. You did not
just guide me academically, you instilled in me the confidence to break
norms and follow this less-conventional career trajectory, for which I
am incredibly passionate.

To my colleagues in the Master\textquotesingle s in Medical Informatics
program, your camaraderie has been invaluable. The diversity and
multi-disciplinary nature of our cohort have not only expanded my
horizons but also deeply enriched my perspective.

A special note of appreciation goes to my colleagues at MTG. Your
innovative spirit and commitment to excellence have contributed greatly
to my professional growth and have continually inspired me to strive for
the best.

Finally, my heartfelt thanks go to the entire team at MEDCIDS. Your warm
welcome and consistent support have been instrumental in shaping me both
academically and as a human being.
