\chapter{Conclusion}\label{conclusion}

This work has explored the specificities of data visualization in
healthcare research, with a particular focus on big datasets and
described the development of a data visualization automation tool.

The original contribution of this work lies in the development of a
specialized data visualization system designed to meet the specific
needs of academic and healthcare settings. While it currently operates
in a local environment, it offers a modular architecture that is ripe
for future expansion and integration into cloud-based platforms.

The system demonstrated its ability to efficiently create and update
complex visualizations, such as Forest Plots, offering substantial
advantages in terms of time and resource efficiency.

Importantly, the development process served as an applied case study in
employing a range of software development methodologies and best
practices, offering significant learning experiences that can inform
future work in this domain.

Several limitations were identified, setting the stage for future
development that could focus on expanding the types of visualizations
supported, increasing scalability, and offering more versatile
deployment options.

Ultimately, the insights gained through this work affirm the power of
data visualization as a critical tool for data interpretation and
decision-making in healthcare research. As this field continues to
evolve, it is anticipated that the integration of specialized tools,
coupled with advancements in software engineering practices, will
further amplify the capabilities of data visualization to serve the
complex needs of healthcare research and beyond.

As a final note, it is worth mentioning that the tool developed through
this work will be actively leveraged in our scientific communication
processes, particularly in the context of real-world evidence. This
incorporation not only adds a practical dimension to the academic
contributions of this research but also paves the way for a sustained
impact on healthcare research and outcomes.
